```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(obisindicators)
library(dplyr)
library(h3)
library(sf)
library(arrow)
library(magick)
library(ggplot2)
library(jsonlite)
library(rgdal)
```

# Create a few helper functions

## Create a function to download a new OBIS snapshot, if you don't have it already
```{r function}

download_obis_snapshot <- function() {
  jsons <- data.frame(jsonlite::fromJSON("https://api.obis.org/export")$results)
  
  parquets <- dplyr::filter(jsons, type == "parquet")
  
  s3_path <- dplyr::filter(parquets, created == max(parquets$created))$s3path
  
  url_parquet = sprintf('https://obis-datasets.ams3.digitaloceanspaces.com/%s',
                        s3_path)
  
  dir_data <- here::here("~/GitProjects")
  
  file_parquet <- file.path(dir_data, basename(url_parquet))
  
  stopifnot(dir.exists(dir_data))
  
  if (!file.exists(file_parquet)) {
    
    print(sprintf(
      "%s does not exist! Downloading from %s",
      file_parquet,
      url_parquet
    ))
    
    download.file(url_parquet, file_parquet, mode = "wb")
    
  } else {
    
    print(sprintf("%s already exists!", file_parquet))
    
  }
  
  return(file_parquet)
  
}
```


## Create function to make grid, calculate metrics for different resolution grid sizes

```{r function}
h3_indicators <- function(occ, resolution = 9) {
  
  ## Compute the indicator on the grid resolution of interest for the points in the identified polygons
  # return h3 cell index for occurrences in polygon
  
  occ_h3 <- occ %>%
    mutate(cell = h3::geo_to_h3(data.frame(decimalLatitude, decimalLongitude), res = RES))
  
  # group by cell index and compute indicators
  
  idx <- obisindicators::calc_indicators(occ_h3)
  
  # convert hexagon ids to spatial features
  # NOTE: DATELINEOFFSET is inv proportional to hex_res b/c we need to look
  #       further from the dateline as hex sizes get bigger.
  
  dl_offset <- 60  # 60 is enough for hex_res >= 1. res 0 is weird; don't use it.
  
  hex_sf <- purrr::map_df(idx$cell, h3::h3_to_geo_boundary_sf) %>%
    sf::st_wrap_dateline(c(
      "WRAPDATELINE=YES",
      glue::glue("DATELINEOFFSET={dl_offset}")
    )) %>%
    dplyr::mutate(hexid = idx$cell)
  
  # merge geometry into indicator table
  
  grid <- hex_sf %>%
    inner_join(idx, by = c("hexid" = "cell"))
  
  return(grid)
}
```


# Function to get OBIS records from snapshot
```{r function}
open_parquet_file <- function(filepath){
  
  occ_all <- arrow::open_dataset(filepath)

  occ <- occ_all %>%
    group_by(
      decimalLongitude, decimalLatitude, species) %>%  # remove duplicate rows, add date_year if testing.
    filter(!is.na(species))  %>%
    summarize(
      records = n(),
      .groups = "drop") %>%
    collect()

}
```

# Lets do some work!

## Get the OBIS records
```{r }

file <- download_obis_snapshot()

occ <- open_parquet_file(file)
    
```

# Identify US waters
```{r}
path <- "data/US_Waters_2024_WGS84/US_Waters_2024_WGS84.shp"

tt <- read_sf(path, crs=4326) %>%
  st_transform(27572)

# ggplot() +
#   geom_sf(data=tt)
```
## Subset to generic "u.s. waters" box to reduce amount of processing for next step 
If needed, reduce by date range too.

```{r}
# 
# date_beg <- 1970
# date_end <- date_beg+1
# 
# 
occ_box <- occ %>%
  filter(between(decimalLatitude,-20,80.0)) %>%
  filter(between(decimalLongitude,-180.0,-20.0) | between(decimalLongitude,120.0,180.0))
# 
# Additional filtering for testing
# %>%
#   filter(
#       date_year >= date_beg,
#       date_year <= date_end)

# occ_box <- occ
```


## Subset OBIS to US Waters
Take all of the points and only return points that exist in the polygons of interest.

```{r}
occ_sf <- sf::st_as_sf(occ_box,
                       coords = c("decimalLongitude", "decimalLatitude"),
                       crs = 4326) %>%
  st_transform(27572)


# test each point in each polygon (# pts by # polygons) TRUE = point is in polygon
logi_point_in_pol <- as.data.frame(sf::st_intersects(occ_sf, tt, sparse = FALSE))


vect <- logi_point_in_pol %>%
  mutate(anyTRUE = if_any(.cols = contains('V')))

occ_in_poly <- occ_box[as.vector(vect$anyTRUE), ]
```

## Create the indicators on H3 grids at specified resolution and write geojson files
Also, tell me how long it took to do this.

```{r}
ptm <- proc.time()

for (RES in 1:6) {
  grid_dec <- h3_indicators(occ_in_poly, resolution = RES)
  
  geojson_string <- geojsonsf::sf_geojson(grid_dec)

  fname <- sprintf("data/indicators_all_res%s.geojson",RES)

  write(x=geojson_string, file=fname)
  
  #proc.time() - ptm
}

# Summary of approx compute times for each RES
# RES 1
#   user  system elapsed 
#  22.32    0.06   22.45 

# RES 2
#   user  system elapsed 
#  34.60    0.11   34.78 

# RES 3
#    user  system elapsed 
#  123.09    0.77  126.94 

# RES 4
#   user  system elapsed 
# 176.58    0.29  178.22

# RES 5
#    user  system elapsed 
#  337.44    0.35  341.98

# RES 6
#    user  system elapsed 
# 1139.92    3.69 1200.70 
```

